Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2019-10-23 manuscript/main.tex

@@

The average distance between yearly populations, as measured by the naive model, was 8.97 $\pm$ 1.35 amino acids (Supplemental Fig.~\ref{sup_fig:distance_of_simulated_populations_between_timepoints}).
As expected, the true fitness model outperformed all other models reducing the distance between populations by 2.16 $\pm$ 1.19 amino acids on average and surpassing the naive model in 32 of 33 (97\%) timepoints (Table~\ref{table_simulated_model_selection}).
With the exception of epitope cross-immunity, all biologically-informed models consistently outperformed the naive model (Fig.~\ref{fig:unadjusted_model_accuracy_and_coefficients_for_simulated_populations}).
LBI was the best of these models, reducing the distance between populations by 1.41 $\pm$ 1.30 amino acids on average.
Indeed, both growth-based models received positive coefficients and outperformed the mechanistic models.
The non-epitope mutations metric received a consistently negative coefficient with an average improvement of 0.71 $\pm$ 0.58 amino acids.
Surprisingly, the composite model of epitope cross-immunity and non-epitope mutations did not perform better than the individual non-epitope mutations model (Supplemental Fig.~\ref{sup_fig:unadjusted_composite_model_accuracy_and_coefficients_for_simulated_populations}).
While epitope cross-immunity was not predictive on its own, the consistently positive coefficient it received in the composite model suggested that it was an informative metric.
From these results, we concluded that our method can accurately estimate the evolution of simulated populations, but that the fitness of samples under our simulated conditions was dominated by purifying selection rather than by positive selection at epitope sites.

>>>>

\rnc{A paragraph explaining the metrics would help.}

@@

The average distance between yearly populations, as measured by the naive model, was 8.97 $\pm$ 1.35 amino acids (Supplemental Fig.~\ref{sup_fig:distance_of_simulated_populations_between_timepoints}).
\rnc{would be good to compare the year-on-year distance to diversity. that would give some sense whether a 2aa reduction is big or small.}
As expected, the true fitness model outperformed all other models reducing the distance between populations by 2.16 $\pm$ 1.19 amino acids on average and surpassing the naive model in 32 of 33 (97\%) timepoints (Table~\ref{table_simulated_model_selection}).
With the exception of epitope cross-immunity, all biologically-informed models consistently outperformed the naive model (Fig.~\ref{fig:unadjusted_model_accuracy_and_coefficients_for_simulated_populations}).
LBI was the best of these models, reducing the distance between populations by 1.41 $\pm$ 1.30 amino acids on average.
Indeed, both growth-based models received positive coefficients and outperformed the mechanistic models.
The non-epitope mutations metric received a consistently negative coefficient with an average improvement of 0.71 $\pm$ 0.58 amino acids.
Surprisingly, the composite model of epitope cross-immunity and non-epitope mutations did not perform better than the individual non-epitope mutations model (Supplemental Fig.~\ref{sup_fig:unadjusted_composite_model_accuracy_and_coefficients_for_simulated_populations}).
While epitope cross-immunity was not predictive on its own, the consistently positive coefficient it received in the composite model suggested that it was an informative metric.
From these results, we concluded that our method can accurately estimate the evolution of simulated populations, but that the fitness of samples under our simulated conditions was dominated by purifying selection rather than by positive selection at epitope sites.
